```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

## Week 1 Assignment

### 1.1 - Population Size 

A key concept in population genetics is **genetic drift**, or the random fluctuation of allele frequencies from generation to generation. In our neutral simulations, genetic drift is the **only** force that causes alleles to become more or less common with each generation. 

The extent that genetic drift can influence allele frequencies is dependent on the population size. We can use our neutral simulation to look further investigate this relationship. Pick a range of population sizes, and run your simulation at least 5 times for each population size. Record the number of fixed mutations for each population size. Briefly comment on how the population size is related to the time to fixation. 

### 1.2 - Selection

So far, we focused on simulations where all mutations were neutral. Change the selection coefficient of your mutations. Try making your mutation slightly favorable. How does the number of fixed mutations change when you add favorable mutations? How low a selection coefficiet do you need for the effect to be visible? 

Do the same with a slightly deleterious mutation. 

### 1.3 - Multiple mutation types; multiple genomic regions

Create two mutation types, one of which is highly (perhaps a selection coefficient around ~0.05) and one of which is deleterious (selection coefficient ~-0.01). 

Create a genomic region with both of these mutation types. Recall that we can use the syntax `c(m1, m2)` to provide multiple items together to an argument. For example, we can initialize a genomic element type with two mutation types at relative frequencies of 0.2 and 0.8 as such: `initializeGenomicElementType("g2", c(m1, m2), c(0.2, 0.8));`. 

Run this a few times and look to see if your deleterious mutation ever reaches fixation. How could you explain this behavior? Does it still happen if you raise your recombination rate higher (e.g. a biologically unrelatstic value of 1e-2)? Why or why not? 

Note that the following code block will print the number of fixed mutations of type `m2`:

```{r, engine='python', eval=FALSE}
2000 late() { 
	print('Number of type 2 fixed mutations:');
	print(sum(sim.substitutions.mutationType == m2));
}
```


