```{r, include = FALSE}
ottrpal::set_knitr_image_path()
``` 

## Multiple Mutations, Elements and Regions 

So far, we have looked at simulations where we only have one mutation type, one genomic element type, and one genomic region. Now, let's look at how to create multiple mutation types and genomic regions.

Here is a simulation that creates three mutation types, each with their own fitness effects

```{r, engine='bash', eval=FALSE}
initialize(){
	initializeMutationRate(1e-7);
	
	initializeMutationType("m1", 1.0, "f", 0.0); // Neutral mutation
	initializeMutationType("m2", 1.0, "f", 0.02); // favorable mutation
	initializeMutationType("m3", 1.0, "f", -0.01); // Deleterious mutation	
	
	initializeGenomicElementType("g1", c(m1, m2, m3), c(50, 1, 10)); 
	initializeGenomicElementType("g2", m1, 1); 
	initializeGenomicElement(g1, 0, 9999);
	initializeGenomicElement(g2, 10000, 15000);
	initializeGenomicElement(g1, 15001, 25000);
	
	initializeRecombinationRate(1e-8);
}

1 early(){
	sim.addSubpop("p1", 500);
}

10000 early()
{
  sim.outputFixedMutations();
} 
```

### Making multiple mutations

We can make multiple mutation types using repeated calls to the `initializeMutationType()` statement. From the script above:

```{r, engine='bash', eval=FALSE}
initializeMutationType("m1", 1.0, "f", 0.0); // Neutral mutation
initializeMutationType("m2", 1.0, "f", 0.02); // favorable mutation
initializeMutationType("m3", 1.0, "f", -0.01); // Deleterious mutation	
```

Notice that each mutation type has its own name. The three mutations represent neutral, favorable, and deleterious mutations, respectively. 














